environmental_data_preperation

### 2018-09-17 src/environmental_data_preperation.Rmd
create unique cell id layer (based on a GLORYS Layer)

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("RNetCDF", "raster", "rgdal")
source("./src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

# read in a netcdf
```{r}
library("RNetCDF") #this will read the netcdf version 1 formats 
library("ncdf4")
ncbr <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
ncbr
```

reproject...
```{r}
canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
library("rgdal")
rp <- projectRaster(ncbr, crs = canada_eq_ar_proj)
rp
```

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
```{r pre-install packages, message=FALSE}
packages <- c("raster", "rgdal") 
source("../src/ipak.R")
ipak(packages)
```

# read in a netcdf
```{r}
library("ncdf4")
ncbr <- brick("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
ncbr
```

# reproject 
There is currently an issue with the projectRaster command in the raster package. Michael Sumners kindly offered a [temporary solution](https://twitter.com/mdsumner/status/1040500830539935744)
```{r}
## use ncdf4 directly
library(ncdf4)
## connection to the file
nc <- nc_open("../data/env/CMEMS_GLOBAL_REANALYSIS_PHY_001_025_salinity_monthly_mean_1993-01.nclonlat.ncdepthtrim.nc")
## desired variable 
v <- ncvar_get(nc, "salinity")
## dimension coordinates
lon <- ncvar_get(nc, "longitude")
lat <- ncvar_get(nc, "latitude")
## a template, one map from the lon, lat, and v[,,1]
template <- raster(list(x = lon, y = lat, z = v[,,1]), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ")

## apply template to brick, and set values to v (but flipped in Y because that's how raster works - lat is increasing in the NetCDF, so we go upside down)
diy <- setValues(brick(template), v[, ncol(v):1, ])

## target
prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs" #note -60 used to be -91. Changed to 'straighten up'.
target <- projectExtent(diy, prj)  ## get a target extent, and we'll set the exact resolution we want
res(target) <- c(25000, 25000)
aea <- projectRaster(diy, target)
plot(aea[[1]]) #the number in brackets refers to the depth number
aea
writeRaster(aea[[1]],'../data/env/aea.tif',overwrite=TRUE,options=c('TFW=YES'))
```

#if projectRaster woudl work properly...

library(raster)
canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=- +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
library("rgdal")
rp <- projectRaster(ncbr, crs = canada_eq_ar_proj)
rp
writeRaster(rp,'../data/env/rp.tif',options=c('TFW=YES')) #this last line jsut saves as a tiff...don't really nead this

----------

#create unique id layer
ok now you want to create a duplicate of the top layer and then give each cell an unique ID

First 
```{r}
aea[[1]]
unique_cell <- (aea[[1]])
```

Because the NA values are currently attributed with some other number and not NA in R (NA is set to -32767 in this particular layer), you need to make sure to exclude the cells which represent NA... so anything under 0.
the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (11,425)
```{r}
cell_val_freq <- freq(unique_cell, value=NULL)
cell_val_freq
```
ok so we have 4693 cells that have NA (or in reality a negative number that represents NA). so remove 4693 from ncell. We know that there are 11424 cells, soo... 11424-4693 = 6732 cells have data.

now give a unique ID to each cell
```{r}
unique_cell[unique_cell > 0] <- 1:6731
writeRaster(unique_cell,'../output/env/unique_cell1.tif',overwrite=TRUE, options=c('TFW=YES')) #this is just to test it in ArcGIS
```
